---
title: "Filters and joins"
author: "Jason"
date: "1/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse) # install.packages("tidyverse")
library(readxl) # install.packages("readxl")
library(here) # install.packages("here")
library(kableExtra) # install.packages("kableExtra")
```

Get some fish data (fish.csv) and kelp data (kelp_fronds.xlsx)

```{r load data}
fish <- read_csv(here("data", "fish.csv"))
kelp_abur <- read_xlsx(here("data", "kelp_fronds.xlsx"))

# add ", sheet = 'sheet_name'" before the last parentheses if there are multiple sheets.
# eading from Excel only pulls in the first worksheet.
```

### Filter by an exact match

Exact match: `==`

Example 1:  only keep observations from `fish` where the common_name variable matches (==) "garibaldi"

```{r Filter exact match}
fish_garibaldi <- fish %>% 
  filter(common_name == "garibaldi")
```

>, <, >=, <=, == as filter match types

Create a subset of fish called fish_over50, starting from fish, where the total_count is greater than OR equal to 50.

```{r fish over 50}
fish_over50 <- fish %>% 
  filter(total_count >= 50)

```

### What if we want to filter by this OR this?

You can use the vertical line `|` for "or"


```{r filtering with multiple criteria and OR}
fish_2sp <- fish %>% 
  filter(common_name == "garibaldi" | common_name == "blacksmith")

# If you're filtering for ultiple matches within the same variable, you can use %in%

fish_3sp <- fish %>% 
  filter(common_name %in% c("garibaldi","blacksmith","black surfperch"))
```

Activity:  subste from fish (fish_gar_2016) that keeps all observations where the year is 2016 OR the common_name is "garibaldi"

```{r two separate conditions with OR}
fish_gar_2016 <- fish %>% 
  filter(common_name == "garibaldi" | year == 2016)
```

### Filter to match this AND that

, or &

```{r two conditions with AND}
aque_2018 <- fish %>% 
  filter(year == 2018, site == "aque")
```

To exclude things: `!=`

### `stringr::str_detect()`

```{r string detect in filter}
fish_black <- fish %>% 
  filter(stringr::str_detect(common_name, pattern = "black"))
```

Subset of data if the common name must include the string "it"?

```{r string detect it in filter}
fish_it <- fish %>% 
  filter(str_detect(common_name, pattern = "it"))
```

### `dplyr::*_join`

```{r full join}
ab_kelpfish <- kelp_abur %>% 
  full_join(fish, by = c("year", "site"))
```

### `left_join()` to keep all from x, and merge matching rows from y (exclude things from y that don't have a match in x)

```{r left join}
kelp_fish_left <- kelp_abur %>% 
  left_join(fish, by = c("year", "site"))

# There is also a right join
```

### `inner_join()` to merge data frames, only keeping observations with a match in both

```{r inner join}
kelp_fish_inner <- kelp_abur %>% 
  inner_join(fish, by = c("year", "site"))
```

### Pipe together multiple steps

- create an object called `my_fish_join`
- starting from `fish`
- filtering to only include observations at site "abur" from the year 2017
- join the kelp_abur data frame to the resulting subset using `left_join()`
- add a new column (`mutate()`) to calculate fish per kelp frond density (counts of fish / total fronds)

```{r piping multiple steps}
my_fish_join <- fish %>% 
  filter(site == "abur" & year == 2017) %>% 
  left_join(kelp_abur, by = c("year", "site")) %>% 
  mutate(fish_per_frond = total_count/total_fronds)

```

Nice HTML table using kableExtra

```{r turn the data frame into a table}
my_fish_join %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```

Activity:  make a chart with x as common_name, and y as fish per frond, geom is column

```{r plot the fish per frond by fish name}
ggplot(data = my_fish_join, aes(x = common_name, y = fish_per_frond)) +
  geom_col(aes(fill = common_name))
```


