---
title: "Pivot Lobsters"
author: "Jason"
date: "1/27/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

Attaching our packages and reading in data

```{r packages_and_data}
library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages("skimr")
library(kableExtra) # install.packages("kableExtra")

# read in lobster data"
lobsters <- read_xlsx(here("data", "lobsters2.xlsx"),
                      skip=4)

```

Viewing our data with skimr

```{r skimming_lobsters}
skimr::skim(lobsters)
# could have run w/o "skimr::", but this provides the package if
# someone is unaware of the package being installed for skim.
```

# Pivot Tables
In R, Pivot tables are: `group_by() %>% summarize()`
`n()` function will count observations

```{r pivot_lobsters}
# group by year
lobsters %>%
  group_by(year) %>% 
  summarize(count_by_year = n())

#group by year and site
lobsters %>%
  group_by(site, year) %>% 
  summarize(count_by_site_year = n())

# ctrl-shft-m is a short-cut for %>%
```

Summarize by multiple variables

```{r pivot_multi_lobsters}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))

# 'mean' and 'sd' will give 'NA' if there are any missing values
# 'na.rm = TRUE' will allow it to give the value without the missing data
```

Table formatting with kableExtra::kable()

```{r build_table}
siteyear_summary %>% 
  kableExtra::kable()
```

There are `r nrow(lobsters)` total lobsters included in this report.

## Activity
1. build from our analyses: calculate the median lobster size for each site and year (hint: 'size_mm' and a function called `median()`
2. sync this to github

```{r activity}
median_size <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size = sd(size_mm, na.rm = TRUE),
            median_size_mm = median(size_mm, na.rm = TRUE))

ggplot(data = median_size, aes(x = year, y = median_size_mm, color = site)) +
  geom_line()
ggsave(here("figures", "lobsters-plot.png"))
```

