---
title: "Creating graphs with ggplot2"
author: "Jason"
date: "1/27/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse. 

We will install these packages:

- tidyverse: suite of packages for wrangling & iz (ggplot)
- readxl: to read in excel files (in many ways)
- here: to help with filepaths

```{r setup, message=FALSE}
library(tidyverse) # install.packages("tidyverse")
library(readxl) # install.packages("readxl")
library(here) # install.packages("here")
```

# California National Parks Data

```{r read-data}

ca_np <- read_csv(here("data", "ca_np.csv"))

ci_np <- read_xlsx(here("data", "ci_np.xlsx"))

# can also use read_xls or read_xlsx as appropriate

```


### Make my own first graph!

We're going to make a line plot of year (x-variable) vs visitors (y-variable) in ggplot:

Insert chunk shortcut of Ctrl-Alt-I
```{r ggplot_init}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()

# Update the geom_ to geom_point(), geom_col(), geom_area()
```

Making a base layer for my CI plots:

```{r ggplot_base}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))

gg_base + geom_point()
```

### Now let's start customizing!

- Line / point colors: color
- Change line type: linetype
- Change point style: pch
- Change transparency: alpha

```{r ggplot_customizing}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

### New geom_point with triangles, color, and a different size

```{r}
gg_base +
  geom_point(
    col = "green",
    pch = 2,
    size = 4
  )
```

### Update graph aesthetics based on variables

```{r}
gg_base +
  geom_point(
    aes(
      size = visitors,
      color = year
    ),
    pch = 6
  )
```

### Use gg-themes for overall updates

- theme_light()
= theme_minimal()
- theme_bw()

```{r ggplot_themes}
gg_base +
  geom_area() +
  theme_classic()
```

### Create a multi-series line plot for all CA National Parks

```{r ggplot_multi_series}
ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name)) +
  geom_line(aes(
    color = park_name
  )) +
  theme_minimal()
```

### Split the charts to individual charts by park name

```{r ggplot_facets}
my_np_graph <- ggplot(data = ca_np, aes(x= year, y = visitors)) +
  geom_point(aes(color = park_name), show.legend = FALSE) +
  facet_wrap(~park_name, scales = "free_y") +
  theme_minimal()
```

Export my graph as a .png
```{r ggplot_export}
ggsave(plot = my_np_graph, here("figures", "np_graph.png"), width = 10, height = 7, units = "in")
```

