---
title: "Tidying session"
author: "Jason"
date: "1/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# Attaching packages
library(tidyverse) # install.packages("tidyverse")
library(janitor) # install.packages("janitor")
library(here) # install.packages("here")
library(readxl) # install.packages("readxl")
```

### Read in inverts.xlsx

```{r read_inverts}
inverts <- read_xlsx(here("data", "inverts.xlsx"))
```

### `pivot_longer()`

Use single colon (:) to indicate a series of sequential columns.  eg: A:D
```{r longer_inverts}
inverts_long <- pivot_longer(data = inverts,
                             cols = '2016':'2018',
                             names_to = "year",
                             values_to = "sp_count")

# What if you want some in series and some NOT in series?
# c('2016':'2018', '2020')
```

What if we want to get it back into wide format?

```{r wider_inverts}
inverts_wide <- inverts_long %>% 
  pivot_wider(names_from = common_name,
              values_from = sp_count)
```

### Use `janitor::clean_names()`

```{r clean_wider_inverts}
inverts_wide <- inverts_wide %>% 
  janitor::clean_names()
```

### `tidyr::unite()` to combine information from columns

Example: merge the site code and year into a single column called "site_year"
```{r merge_inverts_long}
inverts_unite <- inverts_long %>% 
  tidyr::unite(col = "site_year",
               c(site, year),
               sep = "_")
```

### Activity:

Creat a new object called 'inverts_moyr' starting from inverts_long that unties the month and year columns into a SINGLE column names "mo_yr", using a slash "/" as the separator ... then try some other fun separators.

```{r combine_month_year}
inverts_moyr <- inverts_long %>% 
  unite(col="mo_yr",
        c(month, year),
        sep = "/")

# use random separators
inverts_moyr2 <- inverts_long %>% 
  unite(col="mo_yr",
        c(month, year),
        sep = "&")

# can merge more than 2 columns by adding the additional column to the "c()" line
```

### `tidyr::separate()`

```{r separating_column}
inverts_sep <- inverts_unite %>% 
  tidyr::separate(col = site_year,
                  into = c("my_site","my_year"),
                  sep = "_")
```

### `stringr` + `mutate()`

Example: using the built-in `trees1 dataset with mutate() to add 5 to all of the tree heights
```{r mutate data}
trees_new <- trees %>% 
  mutate(height_update = Height + 5)
```

### `str_replace()`

Example: replace california with CA
```{r replace string}
ca_abbr <- inverts %>% 
  mutate(common_name =
           str_replace(common_name,
                       pattern = "california", 
                       =replacement = "CA"))
```

Example: In the common_name column from inverts, you want to replace "spiny" with "awesome" and put it in 'awesome_lobsters'
```{r awesome lobsters}
awesome_lobsters <- inverts %>% 
  mutate(common_name =
           str_replace(common_name,
                       pattern = "spiny", 
                       replacement = "awesome"))

```

